### 2019-05-13 partialrda.Rmd
Genus and core

---
title: "Randomforest"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)
```

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(broom)
library(tibble)
library(knitr)
library(phyloseq)
library(microbiome)
library(purrr)
library(parallel)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

Loading phyloseq object

```{r data}
pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
```

Updating phyloseq objects

```{r updating phyloseq objects}
phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
```

```{r transformations}
rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus  <- microbiome::transform(pseq.genus, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus.core  <- microbiome::transform(pseq.genus.core, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()
```

```{r model function}
modelfunction <- function(x, rda.abund, df.meta) {
    covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
                   "BL_USE_RX_C08", "BL_USE_RX_C09")
    fo <- sprintf("rda.abund ~ %s + Condition(%s)", x, paste0(covariates, collapse = "+"))
    model <- rda(as.formula(fo), data = df.meta)
    anova.term <- anova(model, by="term", perm=100)
    data.frame(var = as.character(x),
               beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
               RDA.Conditioned = summary(model)$partial.chi,
               RDA.Constrained = summary(model)$constr.chi,
               RDA.Unconstrained = summary(model)$unconst.chi,
               p.anova = anova.term[["Pr(>F)"]][1],
               fo = fo)
}
```

```{r return parser}
returnparser <- function(ret){
    res <- do.call(rbind, ret) %>%
        mutate(RDA.total =  RDA.Conditioned + RDA.Constrained + RDA.Unconstrained,
               RDA.Conditioned.percent = RDA.Conditioned/RDA.total,
               RDA.Constrained.percent = RDA.Constrained/RDA.total,
               RDA.Unconstrained.percent = RDA.Unconstrained/RDA.total
               )

    res %>%
        dplyr::select(var,
                      beta,
                      p.anova,
                      RDA.Conditioned.percent,
                      RDA.Constrained.percent,
                      RDA.Unconstrained.percent) %>%
        kable()
}
```

## Species

```{r models species}
ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.species,
              df.meta = meta(pseq.species),
              mc.cores = 5)
returnparser(ret.species)
```

## Genus

```{r models genus}
ret.genus <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.genus,
              df.meta = meta(pseq.genus),
              mc.cores = 5)
returnparser(ret.genus)
```

## Genus core

```{r models genus core}
ret.genus.core <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.genus.core,
              df.meta = meta(pseq.genus.core),
              mc.cores = 5)
returnparser(ret.genus.core)
```

```{r save session}
save.image(file = paste0("session/rda.Rdata"))
```