### 2019-05-14 partialrda.Rmd

---
title: "Redundancy analysis for hypertension"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis', cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)
```

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(broom)
library(tibble)
library(knitr)
library(phyloseq)
library(microbiome)
library(purrr)
library(parallel)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

Loading phyloseq object

```{r data}
pseq.genus <- readRDS("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.species <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
```

Updating phyloseq objects

```{r updating phyloseq objects}
phyloseq::sample_data(pseq.genus) <- phyloseq::sample_data(filter.phenotype.data(pseq.genus))
phyloseq::sample_data(pseq.species) <- phyloseq::sample_data(filter.phenotype.data(pseq.species))
pseq.genus.core <- core(pseq.genus, detection = 0.001, prevalence = 0.01)
```

```{r transformations}
rda.abundances.species  <- microbiome::transform(pseq.species, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus  <- microbiome::transform(pseq.genus, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()

rda.abundances.genus.core  <- microbiome::transform(pseq.genus.core, "clr") %>%
    microbiome::abundances() %>%
    t() %>%
    as.matrix()
```
## All models

```{r model function}
modelfunction <- function(x, rda.abund, df.meta, modelstr = "rda.abund ~ %s + Condition(%s)") {
    covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
                   "BL_USE_RX_C08", "BL_USE_RX_C09")
    fo <- sprintf(modelstr, x, paste0(covariates, collapse = "+"))
    model <- rda(as.formula(fo), data = df.meta)
    anova.term <- anova(model, by="term", perm=1000)
    data.frame(var = as.character(x),
               beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
               RDA.Conditioned = summary(model)$partial.chi,
               RDA.Constrained = summary(model)$constr.chi,
               RDA.Unconstrained = summary(model)$unconst.chi,
               p.anova = anova.term[["Pr(>F)"]][1],
               fo = fo)
}
```

```{r return parser}
returnparser <- function(ret){
    res <- do.call(rbind, ret) %>%
        mutate(RDA.total =  RDA.Conditioned + RDA.Constrained + RDA.Unconstrained,
               RDA.Conditioned.percent = RDA.Conditioned/RDA.total,
               RDA.Constrained.percent = RDA.Constrained/RDA.total,
               RDA.Unconstrained.percent = RDA.Unconstrained/RDA.total
               )

    res %>%
        dplyr::select(var,
                      beta,
                      p.anova,
                      RDA.Conditioned.percent,
                      RDA.Constrained.percent,
                      RDA.Unconstrained.percent) %>%
        kable()
}
```

### Species

```{r models species}
ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.species,
              df.meta = meta(pseq.species),
              mc.cores = 5)
returnparser(ret.species)
```

### Genus

```{r models genus}
ret.genus <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.genus,
              df.meta = meta(pseq.genus),
              mc.cores = 5)
returnparser(ret.genus)
```

### Genus core

```{r models genus core}
ret.genus.core <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.genus.core,
              df.meta = meta(pseq.genus.core),
              mc.cores = 5)
returnparser(ret.genus.core)
```

## Example model (species, systole)

RDA model

```{r rda}
ex.rda <- rda(rda.abundances.species ~ SYSTM + Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X
                                                        + DIAB + BL_USE_RX_C03 + BL_USE_RX_C07 +
                                                        + BL_USE_RX_C08 + BL_USE_RX_C09),
              data = meta(pseq.species))
```

RDA summary

```{r rda summary}
data.frame(Inertia = c(
                   Conditioned = summary(ex.rda)$partial.chi,
                   Constrained = summary(ex.rda)$constr.chi,
                   Unconstrained = summary(ex.rda)$unconst.chi)) %>%
        tibble::rownames_to_column("Variance") %>%
        mutate(Proportion = Inertia/sum(Inertia)) %>%
        mutate_at(2:2, funs(round(., 1))) %>%
        mutate_at(3:3, funs(round(., 5))) %>%
        kable
```

RDA r-squares

```{r rda rsqr}
RsquareAdj(ex.rda)$r.squared %>% kable
```

RDA coefficients

```{r rda coef}
coef(ex.rda) %>% kable
```

Anova

```{r anova}
anova.term <- anova(ex.rda, by="term", perm=1000)
anova.term %>% kable
```

Plot

```{r rda plot}
plot(ex.rda)
```

## Non-partial RDA species

```{r models species}
ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.species,
              df.meta = meta(pseq.species),
              modelstr = "rda.abund ~ %s + %s",
              mc.cores = 5)
returnparser(ret.species)
```

```{r save session}
save.image(file = paste0("session/rda.Rdata"))
```