### 2019-08-28 articletwo-sensitivityanalysis.Rmd
Smaller number of permutations

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
```

# Import previous session

```{r Import previous session}
load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
```

## Variables

```{r my variables}
var.CL.opt <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB", "BP_TREAT")
```

# Sensitivity analysis

```{r sensitivity analysis bp treat}
alphadiversity.askbp <- calculate.alphadiversity(pseq.species,
                                                   list("max" = var.CL.opt,
                                                        "min" = var.CL.min))

betadiversity.askbp <- calculate.betadiversity(pseq = pseq.species,
                                                matrix = bray.dist.m.species,
                                                vars = list("max" = var.CL.opt,
                                                            "min" = var.CL.min),
                                               npermutations = 999)

diversities.helthy <- mergediversities(alphadiversity.askbp,
                                       betadiversity.askbp,
                                       names.dset = names.dset)

plot.diversities(diversities.askbp)
```

Subjects with diabetes, cancer, kidney disease, and coronary disease are dropped.

```{r sensitivity analysis only healthy}
pseq.species.healthy <- subset_samples(pseq.species,
                                       PREVAL_DIAB == 0 &
                                       PREVAL_CHD == 0 &
                                       PREVAL_CR_ANYCANC == 0)

bray.dist.m.species.healthy <- bray.dist.m.species[pseq.species.healthy %>% meta %>% rownames,
                                                   pseq.species.healthy %>% meta %>% rownames]

alphadiversity.healthy <- calculate.alphadiversity(pseq.species.healthy,
                                                   list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                        "min" = var.CL.min))

betadiversity.healthy <- calculate.betadiversity(pseq = pseq.species.healthy,
                                                matrix = bray.dist.m.species.healthy,
                                                vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                            "min" = var.CL.min),
                                                npermutations = 999)

diversities.helthy <- mergediversities(alphadiversity.healthy,
                                       betadiversity.healthy,
                                       names.dset = names.dset)

plot.diversities(diversities.healthy)
```

## BMI

Normal weigth

```{r sensitivity analysis normal weight}
pseq.species.normalweight <- subset_samples(pseq.species, BMI > 18.5 & BMI <= 25)

bray.dist.m.species.normalweight <-
    bray.dist.m.species[pseq.species.normalweight %>% meta %>% rownames,
                        pseq.species.normalweight %>% meta %>% rownames]

alphadiversity.normalweight <- calculate.alphadiversity(pseq.species.normalweight,
                                                   list("max" = var.CL[var.CL != "BMI"],
                                                        "min" = var.CL.min))

betadiversity.normalweight <- calculate.betadiversity(pseq = pseq.species.normalweight,
                                                matrix = bray.dist.m.species.normalweight,
                                                vars = list("max" = var.CL[var.CL != "BMI"],
                                                            "min" = var.CL.min),
                                                npermutations = 999)

diversities.helthy <- mergediversities(alphadiversity.normalweight,
                                       betadiversity.normalweight,
                                       names.dset = names.dset)

plot.diversities(diversities.normalweight)
```

Obese

```{r sensitivity analysis obese}
pseq.species.obese <- subset_samples(pseq.species, BMI > 30 & BMI <= 40)

bray.dist.m.species.obese <-
    bray.dist.m.species[pseq.species.obese %>% meta %>% rownames,
                        pseq.species.obese %>% meta %>% rownames]

alphadiversity.obese <- calculate.alphadiversity(pseq.species.obese,
                                                   list("max" = var.CL[var.CL != "BMI"],
                                                        "min" = var.CL.min))

betadiversity.obese <- calculate.betadiversity(pseq = pseq.species.obese,
                                                matrix = bray.dist.m.species.obese,
                                                vars = list("max" = var.CL[var.CL != "BMI"],
                                                            "min" = var.CL.min),
                                               npermutations = 999)

diversities.helthy <- mergediversities(alphadiversity.obese,
                                       betadiversity.obese,
                                       names.dset = names.dset)

plot.diversities(diversities.obese)
```